Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-04-08 demystify.tex
updates

You \emph{do not} need to do a PhD in CS to work in your area of interest. For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.

However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty. A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenured or tenure-track faculty in CS (\autoref{sec:faculty-types}). 
For example, a prof. in Stats or Math dept might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.  Similarly, a prof. in CS cannot serve as a formal adviser for a PhD student in a Math or Stats.
If in doubt, check with the CS department for their requirements.

For this specific reason,  CSRankings includes only tenure line faculty who can advise CS PhD students. I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students. \autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.

>>>>

You \emph{do not} need to do a PhD in CS to work in your area of interest. For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, both of which have faculty with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in a CS department.

However, if your goal is a PhD in CS, then you likely need to be in the CS dept \emph{and} advised by a CS faculty. A non-CS faculty can serve in the PhD dissertation committee (common) or \emph{co-advise} (somewhat rare), but your main PhD adviser will likely be a tenure-line faculty in CS (\autoref{sec:faculty-types}). 
For example, a prof. in Stats or Math dept might be able to serve as a co-adviser, but not as a sole adviser of a student in a CS PhD program.  Similarly, a prof. in CS cannot be the sole adviser and graduate a PhD student in a Math or Stats.
If in doubt, check with the CS department for their requirements.

For this specific reason,  CSRankings includes only faculty who can advise CS PhD students. I also have compiled a \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Viet-CS-Profs-US}{list} of Vietnamese faculty who can advise PhD students. \autoref{sec:faculty-types} talks more about who can serve as your PhD adviser.
